---
title: "Functions (cont.): Questions 5, 6, 7 - SOLUTIONS"
output: html_document
date: "2024-11-06"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide", warning = FALSE, message = FALSE)
```

NOTE: The following three questions are identical to those presented in the in-class materials from the previous lecture. However, the solutions provided in this folder have been updated to include more detailed explanations.


## QUESTION 5. Write a function that calculates the mean of a numeric vector.
Do not use any R built-in functions, e.g., do not use length(), sum(), etc.
Then add a check to evaluate if the input vector is numeric
If not, use stop() to stop the code and write a message

Solution using other built-in functions:

```{r}
# define a numeric vector
v <- c(1:10)

# our goal
mean(v)
```

```{r}
# compute the mean without the mean() function, but with other built-in functions 
# mean formula: sum of all numbers / count of all numbers
# applied to v: 55 / 10

total_values <- length(v)
result <- sum(v) / total_values
result
```

```{r}
# once code works, put it into a function

my_mean <- function(vector) {
  total_values <- length(vector)
  result <- sum(vector) /total_values
  return(list(total_values, result))
}

my_mean(v)

output <- my_mean(v)

values <- output[[1]]
mean <- output[[2]]

# test this out with another vector 
```

Solution without using other built-in functions:

```{r}
# take our numeric vector v defined above and recall our goal
# mean formula: sum of all numbers / count of all numbers
# applied to v: 55 / 10

v
sum_elements <- 1 + 2 + 3 + 4 + 5 + 6 + 7 + 8 + 9 + 10
count_elements <- length(v)
sum_elements / count_elements
```

```{r}
# compute the mean without any built-in functions

# initialize variables
sum_elements <- 0
count_elements <- 0

# calculate sum of elements in vector (looping over indexes)
  for (i in seq_along(v)) {
    sum_elements <- sum_elements + v[i]
  }
  
  # calculate length of vector 
  for (v in seq_along(v)) {
    count_elements <- count_elements + 1 
  }
  
  # calculate mean
  mean_value <- sum_elements / count_elements
  mean_value
```


```{r}
# once code works, put it into a function
# notice we can put everything into one for loop

mean_vect <- function(vector) {
  
  # initialize variables
  sum_elements <- 0
  count_elements <- 0
  
  # calculate sum and count in one loop
  for (i in seq_along(vector)) {
    sum_elements <- sum_elements + vector[i]
    count_elements <- count_elements + 1
  }
  
  # calculate the mean
  mean_value <- sum_elements / count_elements
  return(mean_value)
}

# call the function on different vectors
mean(v)
mean_vect(c(2,4,6))

```

Add a check to evaluate if vector is numeric, and if not `stop()` the function:

```{r}

mean_vect <- function(vector) {
  
  if (!is.numeric(vector)) {
    stop("Input vector must be numeric!")
  }
  
  sum_elements <- 0
  count_elements <- 0
  
  for (i in seq_along(vector)) {
    sum_elements <- sum_elements + vector[i]
    count_elements <- count_elements + 1
  }
  return(sum_elements / count_elements)
}

# call the function on different vectors
mean(v)
mean_vect(c(2,4,6))
#mean_vect(c(2,"4",6))
```


## QUESTION 6. Add documentation to the function you wrote in question 4 or in 5.

```{r}
mean_vect <- function(vector) {
  
  # Calculate mean of a numeric vector without using built-in functions 
  # Args:
  #   vector: input data is a numeric vector
  # Returns:
  #   single scalar: mean of a numeric vector as a single scalar
  
  if (!is.numeric(vector)) {
    stop("Input vector must be numeric!")
  }
  
  sum_elements <- 0
  count_elements <- 0
  
  for (i in seq_along(vector)) {
    sum_elements <- sum_elements + vector[i]
    count_elements <- count_elements + 1
  }
  return(sum_elements / count_elements)
}
```


## QUESTION 7. Given this plot, rewrite it as a function. Your function should take three arguments like that `(df, var, binwidth = NULL)`

```{r, message = FALSE}
# import library and data
library(tidyverse)
data(diamonds)
```

```{r}
# code to re-write 
diamonds %>% 
  ggplot(aes(x = carat)) + 
  geom_histogram(binwidth = 0.1)
```

```{r}
# solution
hist <- function(df, var, binwidth = NULL) {
  df %>% 
    ggplot(aes(x = {{ var }})) + 
    geom_histogram(binwidth = binwidth)
}

# call the function (both codes work)
hist(diamonds, carat, 0.1)
diamonds %>% hist(carat, 0.3)
```

```{r}
# can also add labels outside function
diamonds %>% hist(carat, 0.1) +
  labs(x = "Size (in carats)", y = "Number of diamonds")
```

