---
title: "Functions (cont.): Questions 5, 6, 7"
output: html_document
date: "2024-11-06"
---

NOTE: The following three questions are identical to those presented in the in-class materials from the previous lecture. However, the solutions provided in this folder have been updated to include more detailed explanations.


## QUESTION 5. Write a function that calculates the mean of a numeric vector.
Do not use any R built-in functions, e.g., do not use length(), sum(), etc.
Then add a check to evaluate if the input vector is numeric
If not, use stop() to stop the code and write a message

```{r}
# define a numeric vector
v <- c(1:10)

# our goal
mean(v)
```

```{r}

```


## QUESTION 6. Add documentation to the function you wrote in question 4 or in 5.

```{r}
mean_vect <- function(vector) {
  
  if (!is.numeric(vector)) {
    stop("Input vector must be numeric!")
  }
  
  sum_elements <- 0
  count_elements <- 0
  
  for (i in seq_along(vector)) {
    sum_elements <- sum_elements + vector[i]
    count_elements <- count_elements + 1
  }
  return(sum_elements / count_elements)
}
```


## QUESTION 7. Given this plot, rewrite it as a function. Your function should take three arguments like that `(df, var, binwidth = NULL)`

```{r, message = FALSE}
# import library and data
library(tidyverse)
data(diamonds)
```

```{r}
# code to re-write 
diamonds %>% 
  ggplot(aes(x = carat)) + 
  geom_histogram(binwidth = 0.1)
```

```{r}

```


